---
title: "BDA - Project brms library test"
author: "Arsi Ikäheimonen"
output:
  pdf_document:
    toc: yes
    toc_depth: 1
  html_document:
    toc: yes
    toc_depth: '1'
    df_print: paged
urlcolor: blue
---

\newpage

Load packages
```{r message=FALSE, warning=FALSE}
library(aaltobda)
library(cmdstanr)
library(brms)
library(ggplot2)
library(corrplot)

SEED <- 614273
```


Load data
```{r}
data <- read.csv('Machine-Learning-with-R-datasets/insurance.csv')
head(data)
```

Check for null values
```{r}
colSums(is.na(data))
```

Some typecasting
```{r}
data$region <- as.factor(data$region)
data$sex <- as.factor(data$sex)
data$smoker <- as.factor(data$smoker)
data$children <- as.integer(data$children)
head(data)
```


Summary statistics of the data
```{r}
summary(data)
```

Plot histogram of the insurance costs
```{r}
ggplot() +
  geom_histogram(aes(data$charges), fill = 'steelblue') +
  labs(title = 'Medical costs billed by health insurance', x='Insurance cost ($)')
```

Plot histogram of the BMI values
```{r}
ggplot() +
  geom_histogram(aes(data$bmi), fill = 'steelblue') +
  labs(title = 'Body mass index', x='BMI (kg/m^2)')
```


Plot scatter plot of the insurance costs with bmi as x-axis.
```{r}
ggplot(data) +
  geom_point(aes(x=bmi, y=charges), size = 1, color = 'steelblue') +
  labs(y = 'Insurance cost ($)', x= 'BMI (kg/m^2)') +
  guides(linetype = "none")
```



Plot scatter plot of the insurance costs with age as x-axis.
```{r}
ggplot(data,aes(x=age,y=charges,col=smoker)) +
  geom_point(size = 1,) +
  labs(y = 'Insurance cost ($)', x= 'Age') +
  guides(linetype = "none")
```
Charges vs region
```{r}
ggplot(data, aes(x=region, y=charges, fill=region)) + 
    geom_boxplot()
```
Charges vs sex
```{r}
ggplot(data, aes(x=sex, y=charges, fill=sex)) + 
    geom_boxplot()
```
Charges vs smoker
```{r}
ggplot(data, aes(x=smoker, y=charges, fill=smoker)) + 
    geom_boxplot()
```
Charges vs children
```{r}
ggplot(data, aes(x=children, y=charges, fill=children, group=children)) + 
    geom_boxplot()
```
Remove region column and label encode sex and smoker columns in order to compute correlation matrix
```{r}
data_clean <- subset(data, select = -c(region))
data_clean$sex <- ifelse(data_clean$sex == 'male',1,0)
data_clean$smoker <- ifelse(data_clean$smoker == 'yes',1,0)
head(data_clean)
```

```{r}
cor_matrix = cor(data_clean)
corrplot(cor_matrix, type = 'lower', order = 'AOE', addCoef.col = 'black', tl.pos = 'd',
         cl.pos = 'n')
```

Basic linear model
```{r}
basic_model = lm(charges~age+sex+bmi+children+smoker+region, data = data) #Create the linear regression
summary(basic_model) #Review the results
```

Baseline model with only Age data as it has the second highest correlation to Charges after smoking.
```{r}
pr = prior(normal(0, 10), class = 'b')

fit_baseline = brm(
  charges ~ age,
  data  = data,
  prior = pr,
  cores = 4
)
```

```{r}
summary(fit_baseline)
```



Bayesian model with brms
```{r}
bayesian_mixed = brm(
  charges ~ age + sex + bmi + children + (1|region) + (1|smoker),
  data  = data,
  prior = pr,
  cores = 4
)
```
Model summary
```{r}
summary(bayesian_mixed, waic=TRUE)
```
STAN code generation
```{r}
make_stancode(charges ~ age + sex + bmi + children + (1|region) + (1|smoker), data=data, family = "gaussian")
```
Conditional effects

```{r}
conditional_effects(bayesian_mixed)
```



