---
output:
 pdf_document:
  toc: no
  keep_tex: true
 html_document:
  toc: yes
  toc_depth: '1'
  df_print: paged
 header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{fontsize=\scriptsize,commandchars=\\\{\}}
---

# Setup

```{r, message = FALSE, warning = FALSE}
# Define a seed for random
SEED <- 614273

# Load packages
library(aaltobda)
library(brms)
library(caret)
library(corrplot)
library(crosstable)
library(dplyr)
library(flextable)
library(ggplot2)
library(ggplotify)
library(ggpubr)
library(lme4)
library(patchwork)

# Define ggplot2 styling
theme_set(theme_classic())
update_geom_defaults("bar", list(fill = "#d3d3d3", color = "#000000"))
single_plot_theme <- theme(text = element_text(size = 20))
multi_plot_theme <- theme(text = element_text(size = 15))
```

# Data Loading

```{r}
data <- read.csv('https://raw.githubusercontent.com/stedy/Machine-Learning-with-R-datasets/master/insurance.csv')
head(data)
data$sex <- as.factor(data$sex)
data$children <- as.integer(data$children)
data$smoker <- as.factor(data$smoker)
data$region <- recode(as.factor(data$region),
               "northeast" = "north",
               "northwest" = "north",
               "southeast" = "south",
               "southwest" = "south")
head(data)
```

# Data Visualization

## Plotting Variables

```{r, message = FALSE, results = 'hide', fig.keep = 'all'}
# Create plots
dist_age <- ggplot(data, aes(x = age)) +
           geom_histogram(breaks = seq(min(data$age), max(data$age), length.out = 11), bins = 10) +
           labs(x = "Age (years)", y = "Frequency") + 
           scale_x_continuous(breaks = as.integer(seq(min(data$age), max(data$age), length.out = 6)))

dist_sex <- ggplot(data, aes(x = sex)) +
           geom_bar(width = 0.5) + labs(x = "Sex", y = "Frequency") +
           geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5) +
           coord_cartesian(ylim = c(0, 750))

dist_bmi <- ggplot(data, aes(x = bmi)) +
           geom_histogram(bins = 30) +
           labs(x = expression(paste('BMI ', bgroup("(", frac("kg", "m"^2), ")"))), y = "Frequency") +
           scale_x_continuous(breaks = as.integer(seq(min(data$bmi), max(data$bmi), length.out = 5)))

dist_children <- ggplot(data, aes(x = children)) +
                geom_bar() + labs(x = "Number of children", y = "Frequency") +
                geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5) +
                scale_x_continuous(breaks = min(data$children):max(data$children),
                                   labels = sapply(min(data$children):max(data$children), function(x) toString(x))) +
                coord_cartesian(ylim = c(0, 650)) 

dist_smoker <- ggplot(data, aes(x = smoker)) +
              geom_bar(width = 0.5) + labs(x = "Smoker", y = "Frequency") +
              geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5) +
              coord_cartesian(ylim = c(0, 1200))

dist_region <- ggplot(data, aes(x = region)) +
              geom_bar(width = 0.5) +
              labs(x = "Region (USA)", y = "Frequency") +
              geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5) +
              coord_cartesian(ylim = c(0, 750)) 

dist_charges <- ggplot(data, aes(x = charges)) +
               geom_histogram(bins = 30) +
               labs(x = "Price ($)", y = "Frequency")

dist_all <- (dist_age + dist_sex) /
            (dist_bmi + dist_region) /
            (dist_children + dist_smoker) /
            dist_charges +
            plot_annotation(tag_levels = 'A') +
            plot_layout(heights = c(1, 1, 1, 1.5)) +
            multi_plot_theme

# Plot all
dist_all

# Save figures
if (!file.exists("img")) dir.create("img")
if (!file.exists("img/dist_plots")) dir.create("img/dist_plots")
ggsave("img/dist_plots/dist_age.pdf", dist_age + single_plot_theme)
ggsave("img/dist_plots/dist_sex.pdf", dist_sex + single_plot_theme)
ggsave("img/dist_plots/dist_bmi.pdf", dist_bmi + single_plot_theme)
ggsave("img/dist_plots/dist_children.pdf", dist_children + single_plot_theme)
ggsave("img/dist_plots/dist_smoker.pdf", dist_smoker + single_plot_theme)
ggsave("img/dist_plots/dist_region.pdf", dist_region + single_plot_theme)
ggsave("img/dist_plots/dist_charge.pdf", dist_charges + single_plot_theme)
ggsave("img/dist_plots/dist_all.pdf", dist_all, width = 8, height = 12)
```

## Data Correlation

```{r, results = 'hide', fig.keep = 'all'}
# Create numeric only data
data_clean <- data
data_clean$sex <- ifelse(data_clean$sex == 'male', 1, 0)
data_clean$smoker <- ifelse(data_clean$smoker == 'yes', 1, 0)
data_clean$region <- ifelse(data_clean$region == 'north', 1, 0)

# Create correlation matrix
cor_matrix <- cor(data_clean)

# Plot correlation matrix
corrplot(cor_matrix, type = 'lower', order = 'AOE', addCoef.col = 'black', tl.pos = 'd', cl.pos = 'n')

# Save figure
if (!file.exists("img")) dir.create("img")
pdf("img/corr_matrix.pdf")
corrplot(cor_matrix, type = 'lower', order = 'AOE', addCoef.col = 'black', tl.pos = 'd', cl.pos = 'n')
dev.off()
```

## Data crosstable

```{r}
# Plot crosstable
crosstable <- crosstable(data, c(sex, region), by = smoker)
crosstable
```

# Pre-processing

## Scale data

```{r}
# Define helper functions
standardize <- function(values) (values - mean(values)) / sd(values)
de_standardize <- function(values, sd_values, mean_values) values * sd_values + mean_values

# Scale the data
data$scaled_charges <- standardize(data$charges)
data$scaled_age <- standardize(data$age)
data$scaled_bmi <- standardize(data$bmi)
data$scaled_children <- standardize(data$children)
head(data)
```

## Split data

```{r}
inTrain <- createDataPartition(y = data$charges, # Target variable
                               p = 0.75, # Percentage of the data to use for training
                               list = FALSE)

train_data <- data[inTrain,]
test_data <- data[-inTrain,]

# Save datasets
if (!file.exists("data")) dir.create("data")
write.csv(train_data, "data/train_data.csv")
write.csv(test_data, "data/test_data.csv")
```

# Modeling

## Baseline model

```{r}
basic_model <- lmer(scaled_charges ~ scaled_age + scaled_bmi + children + sex + (1|smoker) + (1|region),
               data = train_data)
```

## Definition of priors

```{r}
pr1 <- (prior(normal(0, 1), class = "b", coef = "scaled_age"))

pr2 <- (prior(normal(0, 1), class = "b", coef = "scaled_age") + 
       prior(student_t(3, 0, 1), class = "sd", group = "smoker"))

pr3 <- (prior(normal(0, 1), class = "b", coef = "scaled_age") + 
       prior(normal(0, 1), class = "b", coef = "scaled_bmi") + 
       prior(student_t(3, 0, 1), class = "sd", group = "smoker"))

pr4 <- (prior(normal(0, 1), class = "b", coef = "scaled_age") + 
       prior(normal(0, 1), class = "b", coef = "scaled_bmi") + 
       prior(student_t(3, 0, 1), class = "sd", group = "smoker") + 
       prior(student_t(3, 0, 1), class = "sd", group = "region"))

pr5 <- (prior(normal(0, 1), class = "b", coef = "scaled_age") + 
       prior(normal(0, 1), class = "b", coef = "scaled_bmi") + 
       prior(normal(0, 1), class = "b", coef = "scaled_children") + 
       prior(student_t(3, 0, 1), class = "sd", group = "smoker") + 
       prior(student_t(3, 0, 1), class = "sd", group = "region"))

pr6_1 <- prior(normal(0, 1), class = 'b') 

pr6_2 <- (prior(normal(0, 1), class = "b", coef = "scaled_age") + 
         prior(normal(0, 1), class = "b", coef = "sexmale") + 
         prior(normal(0, 1), class = "b", coef = "scaled_bmi") + 
         prior(normal(0, 1), class = "b", coef = "scaled_children") + 
         prior(student_t(3, 0, 1), class = "sd", group = "smoker") + 
         prior(student_t(3, 0, 1), class = "sd", group = "region"))
```

## Define and train the models
```{r}
config <- data.frame(cores = 8, chains = 4, iter = 4000, warmup = 2000)
if (!file.exists("models")) dir.create("models")
model_1 <- brm(scaled_charges ~ scaled_age,
           prior = pr1,
           file = "models/model_1",
           data = train_data,
           cores = config$cores,
           iter = config$iter,
           warmup = config$warmup)

model_2 <- brm(scaled_charges ~ scaled_age + (1|smoker),
           prior = pr2,
           file = "models/model_2",
           data = train_data,
           cores = config$cores,
           iter = config$iter,
           warmup = config$warmup)

model_3 <- brm(scaled_charges ~ scaled_age + scaled_bmi + (1|smoker),
           prior = pr3,
           file = "models/model_3",
           data = train_data,
           cores = config$cores,
           iter = config$iter,
           warmup = config$warmup)

model_4 <- brm(scaled_charges ~ scaled_age + scaled_bmi + (1|smoker) + (1|region),
           prior = pr4,
           file = "models/model_4",
           data = train_data,
           cores = config$cores,
           iter = config$iter,
           warmup = config$warmup)

model_5 <- brm(scaled_charges ~ scaled_age + scaled_bmi + scaled_children + (1|smoker) + (1|region),
           prior = pr5,
           file = "models/model_5",
           data = train_data,
           cores = config$cores,
           iter = config$iter,
           warmup = config$warmup)

model_6_1 <- brm(scaled_charges ~ scaled_age + sex + scaled_bmi + scaled_children +  smoker + region,
             prior = pr6_1,
             file = "models/model_6_1",
             data = train_data,
             cores = config$cores,
             iter = config$iter,
             warmup = config$warmup)

model_6_2 <- brm(scaled_charges ~ scaled_age + sex + scaled_bmi + scaled_children +  (1|smoker) + (1|region), 
             prior = pr6_2,
             file = "models/model_6_2",
             data = train_data,
             cores = config$cores,
             iter = config$iter,
             warmup = config$warmup)
```

# Convergence diagnostics

## Summaries
```{r, warning=FALSE}
summary(basic_model)
summary(model_1)
summary(model_2)
summary(model_3)
summary(model_4)
summary(model_5)
summary(model_6_1)
summary(model_6_2)
```

## Trace plots

```{r, message = FALSE, results = 'hide', fig.keep = 'all'}
# Define figures
trace_1 <- as.ggplot(plot(model_1, plot = FALSE)[[1]][1]) + ggtitle("1 effect model")
trace_2 <- as.ggplot(plot(model_2, plot = FALSE)[[1]][1]) + ggtitle("2 effect model")
trace_3 <- as.ggplot(plot(model_3, plot = FALSE)[[1]][1]) + ggtitle("3 effect model")
trace_4 <- as.ggplot(plot(model_4, plot = FALSE)[[1]][1]) + ggtitle("4 effect model")
trace_5 <- as.ggplot(plot(model_5, plot = FALSE)[[1]][1]) + ggtitle("5 effect model")
trace_6_1 <- as.ggplot(plot(model_6_1, plot = FALSE)[[1]][1]) + ggtitle("6 effect non-hierarchical model")
trace_6_2 <- as.ggplot(plot(model_6_2, plot = FALSE)[[1]][1]) + ggtitle("6 effect hierarchical model")

# Display plots
trace_1
trace_2
trace_3
trace_4
trace_5
trace_6_1
trace_6_2

# Save plots
if (!file.exists("img")) dir.create("img")
if (!file.exists("img/trace_plots")) dir.create("img/trace_plots")
ggsave("img/trace_plots/trace_1.pdf", trace_1 + single_plot_theme, width = 8.5, height = 11)
ggsave("img/trace_plots/trace_2.pdf", trace_2 + single_plot_theme, width = 8.5, height = 11)
ggsave("img/trace_plots/trace_3.pdf", trace_3 + single_plot_theme, width = 8.5, height = 11)
ggsave("img/trace_plots/trace_4.pdf", trace_4 + single_plot_theme, width = 8.5, height = 11)
ggsave("img/trace_plots/trace_5.pdf", trace_5 + single_plot_theme, width = 8.5, height = 11)
ggsave("img/trace_plots/trace_6_1.pdf", trace_6_1 + single_plot_theme, width = 8.5, height = 11)
ggsave("img/trace_plots/trace_6_2.pdf", trace_6_2 + single_plot_theme, width = 8.5, height = 11)
```

## Leave-One-Out Cross Validation (LOO)

```{r}
# Get LOO
loo_1 <- loo(model_1)
loo_2 <- loo(model_2)
loo_3 <- loo(model_3)
loo_4 <- loo(model_4)
loo_5 <- loo(model_5)
loo_6_1 <- loo(model_6_1)
loo_6_2 <- loo(model_6_2)

# Print LOO
loo_1
loo_2
loo_3
loo_4
loo_5
loo_6_1
loo_6_2

# LOO comparisson
loo_compare(loo_1, loo_2, loo_3, loo_4, loo_5, loo_6_1, loo_6_2)
```

# Posterior predictive checks

```{r, message = FALSE, results = 'hide', fig.keep = 'all'}
# Create plots
pp_1 <- pp_check(model_1) + ggtitle("1 effect model")
pp_2 <- pp_check(model_2) + ggtitle("2 effect model ")
pp_3 <- pp_check(model_3) + ggtitle("3 effect model ")
pp_4 <- pp_check(model_4) + ggtitle("4 effect model ")
pp_5 <- pp_check(model_5) + ggtitle("5 effect model")
pp_6_1 <- pp_check(model_6_1) + ggtitle("6 effect non-hierarchical model")
pp_6_2 <- pp_check(model_6_2) + ggtitle("6 effect hierarchical model")
pp_all <- (pp_1 + pp_2) /
          (pp_3 + pp_4) /
          (pp_5 + pp_6_1) /
          (pp_6_2 + plot_spacer()) +
          plot_layout(guides = 'collect') +
          multi_plot_theme

# Plot all
pp_all

# Save plots
if (!file.exists("img")) dir.create("img")
if (!file.exists("img/pp_plots")) dir.create("img/pp_plots")
ggsave("img/pp_plots/pp_1.pdf", pp_1 + single_plot_theme)
ggsave("img/pp_plots/pp_2.pdf", pp_2 + single_plot_theme)
ggsave("img/pp_plots/pp_3.pdf", pp_3 + single_plot_theme)
ggsave("img/pp_plots/pp_4.pdf", pp_4 + single_plot_theme)
ggsave("img/pp_plots/pp_5.pdf", pp_5 + single_plot_theme)
ggsave("img/pp_plots/pp_6_1.pdf", pp_6_1 + single_plot_theme)
ggsave("img/pp_plots/pp_6_2.pdf", pp_6_2 + single_plot_theme)
ggsave("img/pp_plots/pp_all.pdf", pp_all)
```

# Predictive performance

## Data frame to store performance metrics
```{r}
pred_perf <- data.frame(
    Model = c("lm (baseline)", "1-effect", "2-effect", "3-effect", "4-effect", "5-effect", "6-effect-hierarchical", "6-effect-non-hierarchical")
)

# descaling the data
c_sd <- sd(train_data$charges)
c_mean <- mean(train_data$charges)

# de-scaled values
pp_1 <- de_standardize(colMeans(posterior_predict(model_1, newdata = test_data)), c_sd, c_mean)
pp_2 <- de_standardize(colMeans(posterior_predict(model_2, newdata = test_data)), c_sd, c_mean)
pp_3 <- de_standardize(colMeans(posterior_predict(model_3, newdata = test_data)), c_sd, c_mean)
pp_4 <- de_standardize(colMeans(posterior_predict(model_4, newdata = test_data)), c_sd, c_mean)
pp_5 <- de_standardize(colMeans(posterior_predict(model_5, newdata = test_data)), c_sd, c_mean)
pp_6_1 <- de_standardize(colMeans(posterior_predict(model_6_1, newdata = test_data)), c_sd, c_mean)
pp_6_2 <- de_standardize(colMeans(posterior_predict(model_6_2, newdata = test_data)), c_sd, c_mean)
pp_lm <- de_standardize(predict(basic_model, newdata = test_data), c_sd, c_mean)

# Mean average error
calculate_mae <- function(true, predicted) mean(abs(true-predicted))

mae_lm <- calculate_mae(test_data$charges, (pp_lm))
mae_1 <- calculate_mae(test_data$charges, (pp_1))
mae_2 <- calculate_mae(test_data$charges, (pp_2))
mae_3 <- calculate_mae(test_data$charges, (pp_3))
mae_4 <- calculate_mae(test_data$charges, (pp_4))
mae_5 <- calculate_mae(test_data$charges, (pp_5))
mae_6_1 <- calculate_mae(test_data$charges, (pp_6_1))
mae_6_2 <- calculate_mae(test_data$charges, (pp_6_2))

# Add to dataframe
pred_perf <- cbind(pred_perf, mean_absolute_error = c(mae_lm, mae_1, mae_2, mae_3, mae_4, mae_5, mae_6_1, mae_6_2))

# Root mean squared error
calculate_rmse <- function(true, predicted) sqrt(mean((true - predicted)^2))

rmse_lm <- calculate_rmse(test_data$charges, (pp_lm))
rmse_1 <- calculate_rmse(test_data$charges, (pp_1))
rmse_2 <- calculate_rmse(test_data$charges, (pp_2))
rmse_3 <- calculate_rmse(test_data$charges, (pp_3))
rmse_4 <- calculate_rmse(test_data$charges, (pp_4))
rmse_5 <- calculate_rmse(test_data$charges, (pp_5))
rmse_6_1 <- calculate_rmse(test_data$charges, (pp_6_1))
rmse_6_2 <- calculate_rmse(test_data$charges, (pp_6_2))

# Add to dataframe
pred_perf <- cbind(pred_perf, root_mean_squared_error = c(rmse_lm, rmse_1, rmse_2, rmse_3, rmse_4, rmse_5, rmse_6_1, rmse_6_2))

# Root squared error
calculate_r2 <- function(true, predicted) 1 - sum((true - predicted)^2) / sum((true - mean(true))^2)

r2_lm <- calculate_r2(test_data$charges, (pp_lm))
r2_1 <- calculate_r2(test_data$charges, (pp_1))
r2_2 <- calculate_r2(test_data$charges, (pp_2))
r2_3 <- calculate_r2(test_data$charges, (pp_3))
r2_4 <- calculate_r2(test_data$charges, (pp_4))
r2_5 <- calculate_r2(test_data$charges, (pp_5))
r2_6_1 <- calculate_r2(test_data$charges, (pp_6_1))
r2_6_2 <- calculate_r2(test_data$charges, (pp_6_2))

# Add to dataframe
pred_perf <- cbind(pred_perf, root_squared_error = c(r2_lm, r2_1, r2_2, r2_3, r2_4, r2_5, r2_6_1, r2_6_2))

# Print dataframe
pred_perf
```

```{r, message = FALSE, results = 'hide', fig.keep = 'all'}
# Create plots
post_0_1 <- ggplot(test_data, aes(x = charges)) +
            geom_density() +
            ggtitle("Test data") +
            labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_0_2 <- ggplot() +
            geom_density(aes(x = pp_lm)) +
            ggtitle("Linear Mixed Model") +
            labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_1 <- ggplot() +
          geom_density(aes(x = pp_1)) +
          ggtitle("1 effect model") +
          labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_2 <- ggplot() +
          geom_density(aes(x = pp_2)) +
          ggtitle("2 effect model") + 
          labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_3 <- ggplot() +
          geom_density(aes(x = pp_3)) +
          ggtitle("3 effect model") +
          labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_4 <- ggplot() +
          geom_density(aes(x = pp_4)) +
          ggtitle("4 effect model") +
          labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_5 <- ggplot() +
          geom_density(aes(x = pp_5)) +
          ggtitle("5 effect model") +
          labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_6_1 <- ggplot() +
            geom_density(aes(x = pp_6_1)) +
            ggtitle("6 effect non-hierarchical model") +
            labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_6_2 <- ggplot() +
            geom_density(aes(x = pp_6_2)) +
            ggtitle("6 effect hierarchical model") +
            labs(x = "Charges", y = "Density") +
            coord_cartesian(xlim = c(0, 60000), ylim = c(0, 0.0001))

post_all <-  (post_0_1 + post_0_2 + post_1) /
             (post_2 + post_3 + post_4) /
             (post_5 + post_6_1 + post_6_2) +
             multi_plot_theme
# Plot all
post_all

# Save plots
if (!file.exists("img")) dir.create("img")
if (!file.exists("img/post_plots")) dir.create("img/post_plots")
ggsave("img/post_plots/post_0_1.pdf", post_0_1 + single_plot_theme)
ggsave("img/post_plots/post_0_2.pdf", post_0_2 + single_plot_theme)
ggsave("img/post_plots/post_1.pdf", post_1 + single_plot_theme)
ggsave("img/post_plots/post_2.pdf", post_2 + single_plot_theme)
ggsave("img/post_plots/post_3.pdf", post_3 + single_plot_theme)
ggsave("img/post_plots/post_4.pdf", post_4 + single_plot_theme)
ggsave("img/post_plots/post_5.pdf", post_5 + single_plot_theme)
ggsave("img/post_plots/post_6_1.pdf", post_6_1 + single_plot_theme)
ggsave("img/post_plots/post_6_2.pdf", post_6_2 + single_plot_theme)
ggsave("img/post_plots/post_all.pdf", post_all, width = 8.5, height = 11)
```

```{r}
as_flextable(crosstable(data, c(sex, region), by=smoker))
```